---
title: "Discounting of Probabilistic Food Reinforcement by Pigeons"
subtitle: "Confirmatory R replication of Oliveira et al. (2025)"
author:
  - name: Luís Oliveira
  - name: Leonard Green
  - name: Joel Myerson
  - name: Haoran Wan
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    code-tools: true
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
  freeze: false
engine: knitr
---

## Purpose and scope

This document implements the confirmatory analyses reported in:

> Oliveira, L., Green, L., Myerson, J., & Wan, H. (2025). Discounting of probabilistic food reinforcement by pigeons. *Journal of the Experimental Analysis of Behavior, 124*, e70042. <https://doi.org/10.1002/jeab.70042>

The public data are maintained on the [Open Science Framework](https://osf.io/scwg3/). Download `PigeonPD_Exp1.csv` and `PigeonPD_Exp2.csv` and place them in `data/` at the repository root.

The analysis follows the article's distinction between confirmatory and descriptive observations. Replication conditions are displayed but excluded from model fitting and AUC tests. OSF-provided group means are included in descriptive curve fitting but excluded from subject-level amount-effect models.

## Scientific specification

Experiment 1 models relative subjective value (RSV) as

$$
\operatorname{RSV} = \frac{1}{(1+bX)^s},
$$

where $X$ is the odds against receipt of the larger reinforcer. Experiment 2 models choice between two probabilistic reinforcers as

$$
\operatorname{RSV} =
\left(\frac{1+bX_{\mathrm{smaller}}}{1+bX_{\mathrm{larger}}}\right)^s.
$$

Both analyses use beta likelihoods. The positive priors for $b$ and $s$ are Normal(1, 10), as reported in the article. Subject-specific coefficients are estimated directly; these curve-fitting models are therefore not described as partially pooled or hierarchical.

```{r setup}
required_packages <- c(
  "brms", "cmdstanr", "dplyr", "ggplot2", "glmmTMB", "here",
  "knitr", "posterior", "purrr", "readr", "stringr", "tidyr"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages) > 0) {
  stop(
    "Install the missing R packages before rendering: ",
    paste(missing_packages, collapse = ", ")
  )
}

library(brms)
library(dplyr)
library(ggplot2)
library(purrr)
library(readr)
library(stringr)
library(tidyr)

SEED <- 20250714L
SUBJECT_LEVELS <- c(paste0("P", 41:48), "Mean")
OUTPUT_DIR <- here::here("outputs")
MODEL_DIR <- file.path(OUTPUT_DIR, "models")
FIGURE_DIR <- file.path(OUTPUT_DIR, "figures")
TABLE_DIR <- file.path(OUTPUT_DIR, "tables")
walk(c(MODEL_DIR, FIGURE_DIR, TABLE_DIR), dir.create, recursive = TRUE, showWarnings = FALSE)

theme_set(
  theme_classic(base_size = 12) +
    theme(
      strip.background = element_blank(),
      strip.text = element_text(face = "bold"),
      legend.position = "top"
    )
)
```

## Data import and validation

```{r data}
find_data_dir <- function() {
  requested <- Sys.getenv(
    "PIGEON_DATA_DIR",
    unset = here::here("data")
  )
  candidates <- unique(c(requested, here::here("Analysis")))
  required <- c("PigeonPD_Exp1.csv", "PigeonPD_Exp2.csv")
  valid <- keep(candidates, ~ all(file.exists(file.path(.x, required))))
  if (length(valid) == 0) {
    stop(
      "Data files not found. Download PigeonPD_Exp1.csv and ",
      "PigeonPD_Exp2.csv from https://osf.io/scwg3/ and place them in data/."
    )
  }
  if (normalizePath(valid[[1]]) != normalizePath(here::here("data"), mustWork = FALSE)) {
    warning("Using a legacy local data location: ", valid[[1]])
  }
  valid[[1]]
}

assert_columns <- function(data, required, label) {
  absent <- setdiff(required, names(data))
  if (length(absent) > 0) {
    stop(label, " is missing required columns: ", paste(absent, collapse = ", "))
  }
  invisible(data)
}

assert_beta_outcome <- function(data, label) {
  if (any(!is.finite(data$RSV)) || any(data$RSV <= 0 | data$RSV >= 1)) {
    stop(label, " contains RSV values outside the open interval (0, 1).")
  }
  invisible(data)
}

data_dir <- find_data_dir()
raw_exp1 <- read_csv(file.path(data_dir, "PigeonPD_Exp1.csv"), show_col_types = FALSE)
raw_exp2 <- read_csv(file.path(data_dir, "PigeonPD_Exp2.csv"), show_col_types = FALSE)

assert_columns(raw_exp1, c("Subject", "Amount", "Probability", "SV"), "Experiment 1")
assert_columns(
  raw_exp2,
  c("Subject", "Amount", "Multiplier", "Probability_Sm", "Probability_Lg", "SV"),
  "Experiment 2"
)

exp1 <- raw_exp1 |>
  mutate(
    Subject = factor(Subject, levels = SUBJECT_LEVELS),
    Probability = as.character(Probability),
    Rep = str_ends(Probability, "R"),
    Probability_Lg = as.numeric(str_remove(Probability, "R$")) / 100,
    Probability_Sm = 1,
    OAs_Lg = (1 - Probability_Lg) / Probability_Lg,
    OAs_Sm = 0,
    RSV = SV / Amount,
    Experiment = 1L,
    Multiplier = 1
  ) |>
  select(Subject, Experiment, Amount, Multiplier, Rep, Probability_Sm,
         Probability_Lg, OAs_Sm, OAs_Lg, RSV)

exp2 <- raw_exp2 |>
  mutate(
    Subject = factor(Subject, levels = SUBJECT_LEVELS),
    Multiplier = as.character(Multiplier),
    Rep = str_ends(Multiplier, "R"),
    Multiplier = as.numeric(str_remove(Multiplier, "R$")),
    Probability_Sm = Probability_Sm / 100,
    Probability_Lg = Probability_Lg / 100,
    OAs_Sm = (1 - Probability_Sm) / Probability_Sm,
    OAs_Lg = (1 - Probability_Lg) / Probability_Lg,
    RSV = SV / Amount,
    Experiment = 2L
  ) |>
  select(Subject, Experiment, Amount, Multiplier, Rep, Probability_Sm,
         Probability_Lg, OAs_Sm, OAs_Lg, RSV)

assert_beta_outcome(exp1, "Experiment 1")
assert_beta_outcome(exp2, "Experiment 2")

study_data <- bind_rows(exp1, exp2) |>
  arrange(Experiment, Subject, Amount, Multiplier, OAs_Lg)

if (any(is.na(study_data$Subject))) {
  stop("Unexpected subject identifiers were found in the OSF files.")
}

knitr::kable(
  study_data |>
    count(Experiment, Rep, Subject_type = if_else(Subject == "Mean", "Group mean", "Pigeon")),
  caption = "Rows available after validation."
)
```

## Area under the empirical curve

AUC is calculated with the trapezoidal rule after normalizing the observed odds-against values by the largest value within each subject and condition. No unobserved anchor is added. This reproduces the article's empirical AUC values (for example, P41's 16-pellet Experiment 1 AUC is approximately .125).

```{r auc}
trapezoid_auc <- function(x, y) {
  order_index <- order(x)
  x <- x[order_index] / max(x[order_index])
  y <- y[order_index]
  sum(diff(x) * (head(y, -1) + tail(y, -1)) / 2)
}

auc_data <- study_data |>
  filter(!Rep) |>
  group_by(Subject, Experiment, Amount, Multiplier) |>
  summarise(AUC = trapezoid_auc(OAs_Lg, RSV), .groups = "drop")

stopifnot(abs(
  auc_data |>
    filter(Subject == "P41", Experiment == 1, Amount == 16) |>
    pull(AUC) - 0.125
) < 0.001)

write_csv(auc_data, file.path(TABLE_DIR, "empirical_auc.csv"))
knitr::kable(head(auc_data, 12), digits = 3, caption = "Empirical AUC (first 12 rows).")
```

## Bayesian curve fitting

The model schedule matches the article: four chains, 4,000 iterations per chain, and 2,000 warmup iterations. Cached fits are reused unless the model or data change.

The article specifies priors for $b$ and $s$ but does not report a prior for beta precision. The implementation uses a broad Normal(0, 10) prior on each subject-specific log-precision coefficient and states this explicitly to avoid an undocumented default.

```{r models}
model_priors <- c(
  prior(normal(1, 10), class = "b", nlpar = "b", lb = 0),
  prior(normal(1, 10), class = "b", nlpar = "s", lb = 0),
  prior(normal(0, 10), class = "b", dpar = "phi")
)

formula_exp1 <- bf(
  RSV ~ log((1 / (1 + b * OAs_Lg)^s) /
              (1 - (1 / (1 + b * OAs_Lg)^s))),
  b + s ~ 0 + Subject,
  phi ~ 0 + Subject,
  nl = TRUE
)

formula_exp2 <- bf(
  RSV ~ log(
    (((1 + b * OAs_Sm)^s) / ((1 + b * OAs_Lg)^s)) /
      (1 - (((1 + b * OAs_Sm)^s) / ((1 + b * OAs_Lg)^s)))
  ),
  b + s ~ 0 + Subject,
  phi ~ 0 + Subject,
  nl = TRUE
)

fit_condition <- function(data, formula, cache_name) {
  brm(
    formula = formula,
    data = droplevels(data),
    family = Beta(link = "logit", link_phi = "log"),
    prior = model_priors,
    chains = 4,
    cores = min(4, parallel::detectCores(logical = FALSE)),
    iter = 4000,
    warmup = 2000,
    seed = SEED,
    backend = "cmdstanr",
    control = list(adapt_delta = 0.99, max_treedepth = 15),
    save_pars = save_pars(all = TRUE),
    file = file.path(MODEL_DIR, cache_name),
    file_refit = "on_change",
    refresh = 250
  )
}

confirmatory <- study_data |> filter(!Rep)

fits_exp1 <- list(
  `16` = fit_condition(
    filter(confirmatory, Experiment == 1, Amount == 16),
    formula_exp1,
    "exp1_16"
  ),
  `32` = fit_condition(
    filter(confirmatory, Experiment == 1, Amount == 32),
    formula_exp1,
    "exp1_32"
  )
)

exp2_conditions <- tidyr::crossing(Amount = c(16, 32), Multiplier = c(1, .75, .25))
fits_exp2 <- pmap(
  exp2_conditions,
  function(Amount, Multiplier) {
    fit_condition(
      filter(
        confirmatory,
        Experiment == 2,
        .data$Amount == .env$Amount,
        .data$Multiplier == .env$Multiplier
      ),
      formula_exp2,
      sprintf("exp2_%s_m%s", Amount, format(Multiplier, trim = TRUE))
    )
  }
)
names(fits_exp2) <- sprintf(
  "%s_m%s",
  exp2_conditions$Amount,
  format(exp2_conditions$Multiplier, trim = TRUE)
)
```

### Diagnostics

```{r diagnostics}
diagnose_fit <- function(fit, model) {
  draws <- posterior::as_draws_array(fit)
  draw_summary <- posterior::summarise_draws(draws, "rhat", "ess_bulk") |>
    as_tibble() |>
    filter(str_detect(variable, "^b_(b|s)_Subject"))
  sampler <- nuts_params(fit)
  tibble(
    model = model,
    max_rhat = max(draw_summary$rhat),
    min_ess_bulk = min(draw_summary$ess_bulk),
    divergences = sum(sampler$Parameter == "divergent__" & sampler$Value == 1)
  )
}

diagnostics <- bind_rows(
  imap_dfr(fits_exp1, ~ diagnose_fit(.x, paste0("Experiment 1, ", .y, " pellets"))),
  imap_dfr(fits_exp2, ~ diagnose_fit(.x, paste0("Experiment 2, ", .y)))
)

write_csv(diagnostics, file.path(TABLE_DIR, "diagnostics.csv"))
knitr::kable(diagnostics, digits = 3)

if (any(diagnostics$max_rhat > 1.01) || any(diagnostics$divergences > 0)) {
  warning("At least one model requires diagnostic review before interpretation.")
}
```

### Posterior summaries and variance accounted for

Parameter-direction probabilities follow the article: $P(b > 1)$ for $b$ and $P(s < 1)$ for $s$. Values above .975 or below .025 were treated as statistically significant in the publication.

```{r summaries}
parameter_summary <- function(fit, experiment, amount, multiplier = 1) {
  draws <- posterior::as_draws_df(fit) |>
    as_tibble() |>
    select(matches("^b_(b|s)_Subject")) |>
    pivot_longer(everything(), names_to = "term", values_to = "value") |>
    mutate(
      Parameter = if_else(str_starts(term, "b_b_"), "b", "s"),
      Subject = str_remove(term, "^b_(b|s)_Subject")
    )

  draws |>
    group_by(Subject, Parameter) |>
    summarise(
      Estimate = median(value),
      Posterior_SD = sd(value),
      pd = if_else(first(Parameter) == "b", mean(value > 1), mean(value < 1)),
      .groups = "drop"
    ) |>
    mutate(Experiment = experiment, Amount = amount, Multiplier = multiplier)
}

variance_accounted_for <- function(fit, experiment, amount, multiplier = 1) {
  observed <- fit$data
  prediction <- apply(posterior_epred(fit), 2, median)
  observed |>
    mutate(prediction = prediction) |>
    group_by(Subject) |>
    summarise(
      R2 = 1 - sum((RSV - prediction)^2) / sum((RSV - mean(RSV))^2),
      .groups = "drop"
    ) |>
    mutate(Experiment = experiment, Amount = amount, Multiplier = multiplier)
}

params_exp1 <- imap_dfr(
  fits_exp1,
  ~ parameter_summary(.x, experiment = 1, amount = as.numeric(.y))
)
r2_exp1 <- imap_dfr(
  fits_exp1,
  ~ variance_accounted_for(.x, experiment = 1, amount = as.numeric(.y))
)

params_exp2 <- pmap_dfr(
  c(exp2_conditions, list(fit = fits_exp2)),
  ~ parameter_summary(..3, experiment = 2, amount = ..1, multiplier = ..2)
)
r2_exp2 <- pmap_dfr(
  c(exp2_conditions, list(fit = fits_exp2)),
  ~ variance_accounted_for(..3, experiment = 2, amount = ..1, multiplier = ..2)
)

parameter_table <- bind_rows(params_exp1, params_exp2)
r2_table <- bind_rows(r2_exp1, r2_exp2)
write_csv(parameter_table, file.path(TABLE_DIR, "posterior_parameters.csv"))
write_csv(r2_table, file.path(TABLE_DIR, "variance_accounted_for.csv"))

knitr::kable(
  r2_exp1 |> filter(Subject != "Mean") |> group_by(Amount) |>
    summarise(Median_R2 = median(R2), Mean_R2 = mean(R2)),
  digits = 3,
  caption = "Experiment 1 model fit across individual pigeons."
)
```

Published reference values are median $R^2=.892$ and $.791$ for the 16- and 32-pellet phases of Experiment 1. Across Experiment 2 conditions, the published median was $.895$ and the mean was $.804$. Small differences are expected from Monte Carlo sampling and software versions; material discrepancies require investigation.

## Model-based discounting curves

Curves show posterior-median expected RSV. Filled points identify confirmatory observations; open points identify replication conditions. Experiment 2 uses the article's relative odds-against definition: the absolute difference between the two odds-against values multiplied by the condition multiplier.

```{r discounting-curves}
#| fig-width: 8
#| fig-height: 8

prediction_curve <- function(fit, experiment, amount, multiplier = 1) {
  model_data <- fit$data
  grid <- crossing(
    Subject = unique(model_data$Subject),
    OAs_Lg = seq(min(model_data$OAs_Lg), max(model_data$OAs_Lg), length.out = 120)
  ) |>
    mutate(
      Subject = factor(Subject, levels = levels(model_data$Subject)),
      OAs_Sm = if (experiment == 1) 0 else unique(model_data$OAs_Sm),
      Experiment = experiment,
      Amount = amount,
      Multiplier = multiplier
    )
  grid$RSV <- apply(posterior_epred(fit, newdata = grid), 2, median)
  grid
}

curves_exp1 <- imap_dfr(
  fits_exp1,
  ~ prediction_curve(.x, experiment = 1, amount = as.numeric(.y))
)
curves_exp2 <- pmap_dfr(
  c(exp2_conditions, list(fit = fits_exp2)),
  ~ prediction_curve(..3, experiment = 2, amount = ..1, multiplier = ..2)
)

observed_plot <- study_data |>
  mutate(
    Amount_label = factor(
      Amount,
      levels = c(32, 16),
      labels = c("32 pellets", "16 pellets")
    ),
    Relative_OAs = abs(OAs_Lg - OAs_Sm) * Multiplier
  )

curve_plot_data <- bind_rows(curves_exp1, curves_exp2) |>
  mutate(
    Amount_label = factor(
      Amount,
      levels = c(32, 16),
      labels = c("32 pellets", "16 pellets")
    ),
    Relative_OAs = abs(OAs_Lg - OAs_Sm) * Multiplier
  )

amount_colors <- c("32 pellets" = "#3568A8", "16 pellets" = "#C66A2B")
amount_lines <- c("32 pellets" = "dashed", "16 pellets" = "solid")
amount_shapes <- c("32 pellets" = 25, "16 pellets" = 24)

plot_curves <- function(experiment, multiplier = 1) {
  observed <- observed_plot |>
    filter(Experiment == experiment, .data$Multiplier == .env$multiplier)
  curves <- curve_plot_data |>
    filter(Experiment == experiment, .data$Multiplier == .env$multiplier)
  x_variable <- if (experiment == 1) "OAs_Lg" else "Relative_OAs"

  ggplot(observed, aes(x = .data[[x_variable]], y = RSV, color = Amount_label)) +
    geom_line(
      data = curves,
      aes(linetype = Amount_label, group = Amount_label),
      linewidth = 0.7
    ) +
    geom_point(
      data = filter(observed, !Rep),
      aes(shape = Amount_label, fill = Amount_label),
      size = 2.2
    ) +
    geom_point(
      data = filter(observed, Rep),
      aes(shape = Amount_label),
      fill = "white",
      size = 2.2,
      show.legend = FALSE
    ) +
    facet_wrap(~ Subject, ncol = 3) +
    scale_color_manual(values = amount_colors, name = NULL) +
    scale_fill_manual(values = amount_colors, name = NULL) +
    scale_linetype_manual(values = amount_lines, name = NULL) +
    scale_shape_manual(values = amount_shapes, name = NULL) +
    scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, .2)) +
    labs(
      x = if (experiment == 1) "Odds against" else "Relative odds against",
      y = "Mean relative subjective value",
      title = if (experiment == 1) NULL else paste("Multiplier", multiplier)
    )
}

figure_1 <- plot_curves(experiment = 1)
figure_3 <- plot_curves(experiment = 2, multiplier = 1)
figure_4 <- plot_curves(experiment = 2, multiplier = .75)
figure_5 <- plot_curves(experiment = 2, multiplier = .25)

figure_1
figure_3
figure_4
figure_5

iwalk(
  list(figure_1 = figure_1, figure_3 = figure_3, figure_4 = figure_4, figure_5 = figure_5),
  ~ ggsave(
    file.path(FIGURE_DIR, paste0(.y, ".pdf")),
    .x,
    width = 8,
    height = 8,
    units = "in"
  )
)
```

## Amount effects on empirical AUC

```{r amount-effects}
auc_subjects <- auc_data |> filter(Subject != "Mean")

auc_model_exp1 <- glmmTMB::glmmTMB(
  AUC ~ factor(Amount) + (1 | Subject),
  family = glmmTMB::beta_family(link = "logit"),
  data = filter(auc_subjects, Experiment == 1)
)

auc_models_exp2 <- auc_subjects |>
  filter(Experiment == 2) |>
  split(.$Multiplier) |>
  map(~ glmmTMB::glmmTMB(
    AUC ~ factor(Amount) + (1 | Subject),
    family = glmmTMB::beta_family(link = "logit"),
    data = .x
  ))

extract_amount_test <- function(model, label) {
  coefficient <- summary(model)$coefficients$cond["factor(Amount)32", ]
  tibble(
    Condition = label,
    Estimate = unname(coefficient["Estimate"]),
    SE = unname(coefficient["Std. Error"]),
    z = unname(coefficient["z value"]),
    p = unname(coefficient["Pr(>|z|)"])
  )
}

amount_tests <- bind_rows(
  extract_amount_test(auc_model_exp1, "Experiment 1"),
  imap_dfr(auc_models_exp2, ~ extract_amount_test(.x, paste("Experiment 2, multiplier", .y)))
)

write_csv(amount_tests, file.path(TABLE_DIR, "auc_amount_effects.csv"))
knitr::kable(amount_tests, digits = 3)
```

The published Experiment 1 contrast was not significant ($p=.37$); all Experiment 2 contrasts were also nonsignificant (all $p>.17$). Interpret the reproduced tests against those reference values and the fitted-model diagnostics.

## Descriptive figure: AUC differences

```{r auc-figure}
#| fig-width: 8
#| fig-height: 4.5

auc_difference <- auc_subjects |>
  select(Subject, Experiment, Multiplier, Amount, AUC) |>
  pivot_wider(names_from = Amount, values_from = AUC, names_prefix = "AUC_") |>
  mutate(
    Difference = AUC_16 - AUC_32,
    Condition = if_else(
      Experiment == 1,
      "Experiment 1",
      paste0("Experiment 2: multiplier ", Multiplier)
    )
  )

auc_plot <- ggplot(auc_difference, aes(Subject, Difference)) +
  geom_hline(yintercept = 0, linetype = 2, color = "grey45") +
  geom_point(size = 2.5, color = "#2C6E49") +
  facet_wrap(~ Condition, scales = "free_x") +
  labs(
    x = "Pigeon",
    y = "AUC difference (16 - 32 pellets)",
    caption = "Positive values indicate steeper discounting of the larger amount."
  )

auc_plot
ggsave(
  file.path(FIGURE_DIR, "auc_amount_differences.pdf"),
  auc_plot,
  width = 8,
  height = 4.5,
  units = "in"
)
```

## Reproducibility record

```{r session-info}
sessionInfo()
cmdstanr::cmdstan_version()
```
